An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
Making top k notation consistent

\btw{The story we discussed: YK show $L_2$ is inconsistent.  Our framework lets us ask: with what discrete loss \emph{is} $L_2$ consistent?  (It also assures us that $L_2$ is consistent for something...) Oh look, $L_2$ embeds something which is $\elltopk$ plus some other term, so (a) we can see very clearly what $L_2$ ``is doing'', and (b) we can see where to look for distributions yielding inconsistency, namely ones for which we actually prefer to report a set with $|r|<k$.}

@@

In Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we show that the set $\U = \{u \in \{0,1\}^n : |u\|_0 \leq k\}$ is always represented among the optimizers of $L'$, meaning for all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$\raft{Jessie, you consistently forget the $p$ (was $u \in \prop{L_2}$)}.
From Lemma~\ref{lem:loss-restrict}\raft{was Lemma~\ref{lem:top-k-surrogate-embeds}}, then, $L'$ must embed $L'|_\U$, which gives us $\ell':\R'\to\reals^\Y_+$ via the natural embedding from sets $\R' = \{r \subseteq [n] : |r| \leq k\}$ to their indicator vectors.
\begin{align}\label{eq:ell-2}
\ell'(r)_y &= \Ind{y \not\in r} + \tfrac 1 k |r \setminus \{y\}|~.
\end{align}
We now see that $\ell'$ is essentially $\elltopk$ (extended to smaller sets) plus an additional cardinality term.

@@

% We now focus on the case where $k=1$; the top-$k$ loss reduces to 0-1 loss in this case.
% To show a set of distributions where $\prop{\ell_2}(p) \neq \prop{\elltopk}(p)$, we claim that $\vec{0}~\in~\prop{\ell_2} \implies \risk{\ell_2}(p) = 1$ for any $p$ with $\max_y p_y \leq 1/2$.
% Additionally $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1 \implies \vec{0} \not\in \prop{\elltopk}(p)$.

% First, $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
% Taking the set $\{y\}$ for any $y \in \Y$, the expected loss $\inprod{p}{\ell_2(\{y\})} = 2 (1-p_y) \geq 1$ by assumption on $p$. 
% %For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r)_y \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
% Therefore, we have that $\vec{0} \in \prop{\ell_2}(p)$ but $\vec{0} \not\in \prop{\elltopk}(p)$ for any $p$ such that $\max_y p_y \leq 1/2$.

%Since we can see that $\ell_2 \neq \elltopk$, we can then see that $L_2$ is not consistent with respect to $\elltopk$.
%Similarly, given $L_4(u,y) := \max\left(\frac{1}{k} \sum_{i=1}^k (1 + u_{\setminus y})_{[i]} - s_y, 0\right)$, we can actually see that $L_4$ also embeds $\ell_2$.
%We will double up on notation and call this discrete loss $\ell_4$ as well, for the sake of matching subscripts of discrete losses to the surrogates that embed them.

@@

\section{Top-$k$ surrogate}
Throughout this section, consider the surrogate and discrete loss $L_2(u)_y~:=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell_2(r)_y = \Ind{y \not\in r} + \frac{|r \setminus \{y\}|}{k}$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

Observe that $L_2$ is a polyhedral loss, since it can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \ones - e_y$ with highest weight.
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L_2}$ is also polyhedral.

\begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$, there is a $u \in \{ u \in \{0,1\}^n : \|u\|_1 \leq k\}$ such that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
\end{lemma}
\begin{proof}[Sketch]
  First, observe that $L_2$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.

  We first claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  Observe that if $u_{[1]} \geq u_{[n]} + 1$, the value of the loss gets ``cropped'' and will be the same by reporting $u_{[1]} = u_{[n]} + 1 = 1$. \jessiet{Flesh this out later.}
  Since $u_{[n]} = 0 \implies u_{[1]} \in [0,1]$, we can then observe there is some $u \in \prop{L_2}$ so that $u \in [0,1]^n$.

  Now, since we can restrict to the closed hypercube, we claim that we can rewrite the expected loss of $L_2$ as 
  \begin{align*}
  	\inprod{p}{L_2(u)} &= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}

@@

  Substituting these values into the expected loss, we now observe a linear objective over the hypercube, which is known to be optimal on some corner of the hypercube, so the rest of $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
  Therefore, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L_2$.
  Moreover, for this optimal $u$, observe $\|u\|_{1} \leq k$ since we have $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.
\end{proof}

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:\R \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{ u \in \{0,1\}^n : \|u\|_1 \leq k \}$.
  Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \varphi(\R)} \inprod{p}{L_2(u)} = \min_{r \in \R}\inprod{p}{\ell_2(r)}$ for all $p \in \simplex$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}

>>>>

\btw{The story we discussed: YK show $L'$ is inconsistent.  Our framework lets us ask: with what discrete loss \emph{is} $L'$ consistent?  (It also assures us that $L'$ is consistent for something...) Oh look, $L'$ embeds something which is $\elltopk$ plus some other term, so (a) we can see very clearly what $L'$ ``is doing'', and (b) we can see where to look for distributions yielding inconsistency, namely ones for which we actually prefer to report a set with $|r|<k$.}

@@

In Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we show that the set $\U = \{u \in \{0,1\}^n : |u\|_0 \leq k\}$ is always represented among the optimizers of $L'$, meaning for all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$\raft{Jessie, you consistently forget the $p$ (was $u \in \prop{L'}$)}.
From Lemma~\ref{lem:loss-restrict}\raft{was Lemma~\ref{lem:top-k-surrogate-embeds}}, then, $L'$ must embed $L'|_\U$, which gives us $\ell':\R'\to\reals^\Y_+$ via the natural embedding from sets $\R' = \{r \subseteq [n] : |r| \leq k\}$ to their indicator vectors.
\begin{align}\label{eq:ell-2}
\ell'(r)_y &= \Ind{y \not\in r} + \tfrac 1 k |r \setminus \{y\}|~.
\end{align}
We now see that $\ell'$ is essentially $\elltopk$ (extended to smaller sets) plus an additional cardinality term.

@@

% We now focus on the case where $k=1$; the top-$k$ loss reduces to 0-1 loss in this case.
% To show a set of distributions where $\prop{\ell'}(p) \neq \prop{\elltopk}(p)$, we claim that $\vec{0}~\in~\prop{\ell'} \implies \risk{\ell'}(p) = 1$ for any $p$ with $\max_y p_y \leq 1/2$.
% Additionally $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1 \implies \vec{0} \not\in \prop{\elltopk}(p)$.

% First, $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
% Taking the set $\{y\}$ for any $y \in \Y$, the expected loss $\inprod{p}{\ell'(\{y\})} = 2 (1-p_y) \geq 1$ by assumption on $p$. 
% %For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell'(r)_y \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
% Therefore, we have that $\vec{0} \in \prop{\ell'}(p)$ but $\vec{0} \not\in \prop{\elltopk}(p)$ for any $p$ such that $\max_y p_y \leq 1/2$.

%Since we can see that $\ell' \neq \elltopk$, we can then see that $L'$ is not consistent with respect to $\elltopk$.
%Similarly, given $L_4(u,y) := \max\left(\frac{1}{k} \sum_{i=1}^k (1 + u_{\setminus y})_{[i]} - s_y, 0\right)$, we can actually see that $L_4$ also embeds $\ell'$.
%We will double up on notation and call this discrete loss $\ell_4$ as well, for the sake of matching subscripts of discrete losses to the surrogates that embed them.

@@

\section{Top-$k$ surrogate}
Throughout this section, consider the surrogate and discrete loss $L'(u)_y~=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell'(r)_y = \Ind{y \not\in r} + \frac{1}{k} |r \setminus \{y\}|$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

\begin{lemma}\label{lem:top-k-polyhedral}
$L'$ is a polyhedral loss.
\end{lemma}
\begin{proof}
Observe $L'$ can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \ones - e_y$, and the max comes from selecting the $u_i$ elements with highest weight.
\end{proof}
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L'}$ is also polyhedral.

\begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$, there is a $u \in \{ u \in \{0,1\}^n : \|u\|_1 \leq k\}$ such that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L'(u)}$.
\end{lemma}
\begin{proof}[Sketch]
  First, observe that $L'$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.

  We first claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  Observe that if $u_{[1]} \geq u_{[n]} + 1$, the value of the loss gets ``cropped'' and will be the same by reporting $u_{[1]} = u_{[n]} + 1 = 1$. \jessiet{Flesh this out later.}
  Since $u_{[n]} = 0 \implies u_{[1]} \in [0,1]$, we can then observe there is some $u \in \prop{L'}$ so that $u \in [0,1]^n$.

  Now, since we can restrict to the closed hypercube, we claim that we can rewrite the expected loss of $L'$ as 
  \begin{align*}
  	\inprod{p}{L'(u)} &= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}

@@

  Substituting these values into the expected loss, we now observe a linear objective over the hypercube, which is known to be optimal on some corner of the hypercube, so the rest of $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
  Therefore, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$.
  Moreover, for this optimal $u$, observe $\|u\|_{1} \leq k$ since we have $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.
\end{proof}

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L':\reals^n \to \reals^\Y_+$ embeds $\ell':\R \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell'(r) = L'(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{ u \in \{0,1\}^n : \|u\|_1 \leq k \}$.
  Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L'(u)} = \min_{u \in \varphi(\R)} \inprod{p}{L'(u)} = \min_{r \in \R}\inprod{p}{\ell'(r)}$ for all $p \in \simplex$, and so we conclude that $L'$ embeds $\ell'$.
\end{proof}
